A Global Genome-wide Scan with Optimal Cutoff Mining for Emerging Biomarkers in Head and Neck Squamous Cell Carcinoma

### 2021-03-05 main.tex
add items from 2021/03/01

*** interpretable or explainable CNN

https://www.cancer.gov/research/areas/diagnosis/artificial-intelligence
% NCI
Aiding the Genomic Characterization of Tumors: 
deep learning could predict commonly mutated genes and mRNA expression from the images
*** A deep learning model to predict RNA-Seq expression of tumours from whole slide images\cite{Schmauch2020}
Preprocessing of whole-slide images. The application of deep-learning algo- rithms to histological data is a challenging problem, particularly due to the high dimensionality of the data (up to 100,000 × 100,000 pixels for a single whole-slide image) and the small size of available datasets. We divided the whole-slide images into squares of 112 × 112 μm (224 × 224 pixels) called “tiles”, and used the Otsu algorithm62 (as implemented in python package skimage) to select only those containing tissue, excluding the white background. We sampled a maximum of 8000 such tiles from each slide. We then extracted 2048 features from those tiles with a 50-layer ResNet63 pretrained on the ImageNet dataset64 (using the Keras implementation), such that a slide could be represented as a 8000 × 2048 matrix.
For the first phase of this work (transcriptome prediction), we accelerated the training of our models through a simple preprocessing step inspired by simple linear iterative clustering (SLIC)65: we used the k-means algorithm (as implemented in python package libkmcuda) to create 100 clusters (supertiles) of tiles on the basis of tile location on the slide, and we averaged the features of the tiles within each cluster. The use of these supertiles reduces the dimensions of a slide to 100 × 2048. The model was first trained on this reduced dataset, with all the TCGA data. Then, for specific organs, fine-tuning was achieved with full-scale data from the organs concerned only.

@@

%discussion: the third limitation
Even there are eighty one ($X_1 ... X_81$) physical, pathological and social features of patients available for survival modelling in the TCGA, 
% since it has comprehensive clinical features, 
such as age, gender, residual tumor, vital status, days\_to\_last\_followup, cancer stage, smoking duration, exposure to alcohol, asbestos, radioactive radon, the TCGA did not collect other holistic features from their participants.
% Input $X_1...X_n$ (e.x. patients features: such as age, gender, gene expression, cancer stage, 
Going for holistic cancer care\cite{Mehta2019}\cite{Iftikhar2021} spiritual and emotional condition is equal important comparing with physical and social status.
%
World Health Organisation has called worldwide for delivering people-centred (or person-centred) care from the physical, emotional, socioeconomic and spiritual perspectives\cite{WHO2015}\cite{Ling-ChengMong2021}.
Core principles can be taken to achieve people-centred and integrated health service delivery. 
First principle is holistic – focusing on physical, socioeconomic, mental and emotional well-being. 
Holistic care: care and see the person as a whole with psychological, social and environmental levels of needs and goals.
%that considers  factors rather than just the symptoms of disease or ill-health.
Second principle is Empowering – supporting people to manage and take responsibility for their own health. 
Empowerment: the process of supporting people and communities to take control of their own health needs resulting, for example, in the uptake of healthier behaviours or the ability to self-manage illnesses.
People-centred care requires that people have the education and support they need to make decisions and participate in their own care.
%Glossary:
%People-centred care: an approach to care that consciously adopts individuals’, carers’, families’ and communities’ perspectives as participants in, and beneficiaries of, trusted health systems that respond to their needs and preferences in humane and holistic ways. People-centred care also requires that people have the education and support they need to make decisions and participate in their own care. It is organized around the health needs and expectations of people rather than diseases (7,8) 
%Person-centred care: care approaches and practices that see the person as a whole with many levels of needs and goals, with these needs coming from their own personal social determinants of health.
%(1. World Health Organization. WHO global strategy on people-centred and integrated health services: interim report. https://apps.who.int/iris/handle/10665/155002 (2015).)

@@

@@

ground truth Y (e.x. patient's survival):\\[0.5cm]
%Y$:\\[0.3cm]
$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n + \epsilon$
gene expression is not an independent $X_mrna$, which is correlated with other $X_body, X_stress, X_inflammation, X_ros$
There is much work to be done in investigating the functional relationships more fully.\cite{Rogers1959}
We wish the TCGA does collect holistic features from their participants in the future. % electric healthcare record, EHR, enabled in holistic way
Deep Patient

@@

>>>>

*** interpretable or explainable CNN
Manning meap book: 不再是黑盒子
https://www.manning.com/books/trust-in-machine-learning
Trust in Machine Learning

*** 必備 explainable interpretable 
By saliency mapping: e.x. diabetes retinopathy, model focused on retinal aterry

https://bmcmedgenomics.biomedcentral.com/articles/10.1186/s12920-019-0624-2
#XAI
Interpretable deep neural network for cancer survival analysis by integrating genomic and clinical data

Jie Hao, 2019
By PyTorch

https://www.deeplearningbook.org/
https://www.cancer.gov/research/areas/diagnosis/artificial-intelligence
% NCI
Aiding the Genomic Characterization of Tumors: 
deep learning could predict commonly mutated genes and mRNA expression from the images
*** A deep learning model to predict RNA-Seq expression of tumours from whole slide images\cite{Schmauch2020}
Preprocessing of whole-slide images. The application of deep-learning algo- rithms to histological data is a challenging problem, particularly due to the high dimensionality of the data (up to 100,000 × 100,000 pixels for a single whole-slide image) and the small size of available datasets. We divided the whole-slide images into squares of 112 × 112 μm (224 × 224 pixels) called “tiles”, and used the Otsu algorithm62 (as implemented in python package skimage) to select only those containing tissue, excluding the white background. We sampled a maximum of 8000 such tiles from each slide. We then extracted 2048 features from those tiles with a 50-layer ResNet63 pretrained on the ImageNet dataset64 (using the Keras implementation), such that a slide could be represented as a 8000 × 2048 matrix.
For the first phase of this work (transcriptome prediction), we accelerated the training of our models through a simple preprocessing step inspired by simple linear iterative clustering (SLIC)65: we used the k-means algorithm (as implemented in python package libkmcuda) to create 100 clusters (supertiles) of tiles on the basis of tile location on the slide, and we averaged the features of the tiles within each cluster. The use of these supertiles reduces the dimensions of a slide to 100 × 2048. The model was first trained on this reduced dataset, with all the TCGA data. Then, for specific organs, fine-tuning was achieved with full-scale data from the organs concerned only.

@@

%discussion: the third limitation
There are eighty one ($X_1 ... X_81$) physical, pathological and social features of patients available for survival modelling in the TCGA, 
% since it has comprehensive clinical features, 
such as age, gender, residual tumor, vital status, days\_to\_last\_followup, cancer stage, smoking duration, exposure to alcohol, asbestos, radioactive radon. However, the TCGA did not collect other holistic features from their participants.
% Input $X_1...X_n$ (e.x. patients features: such as age, gender, gene expression, cancer stage, 
Going for holistic cancer care\cite{Mehta2019}\cite{Iftikhar2021} spiritual and emotional condition is equal important comparing with physical and social status.
%
World Health Organisation has called worldwide for delivering people-centred (or person-centred) care from the physical, emotional, socioeconomic and spiritual perspectives\cite{WHO2015}\cite{Ling-ChengMong2021}.
Core principles can be taken to achieve people-centred and integrated health service delivery. 
First principle is holistic – focusing on physical, socioeconomic, mental and emotional well-being. 
Holistic care: care and see the person as a whole with psychological, social and environmental levels of needs and goals.
%that considers  factors rather than just the symptoms of disease or ill-health.
Second principle is Empowering – supporting people to manage and take responsibility for their own health. 
Empowerment: the process of supporting people and communities to take control of their own health needs resulting, for example, in the uptake of healthier behaviours or the ability to self-manage illnesses.
People-centred care requires that people have the education and support they need to make decisions and participate in their own care.
%Glossary:
%People-centred care: an approach to care that consciously adopts individuals’, carers’, families’ and communities’ perspectives as participants in, and beneficiaries of, trusted health systems that respond to their needs and preferences in humane and holistic ways. People-centred care also requires that people have the education and support they need to make decisions and participate in their own care. It is organized around the health needs and expectations of people rather than diseases (7,8) 
%Person-centred care: care approaches and practices that see the person as a whole with many levels of needs and goals, with these needs coming from their own personal social determinants of health.
%(1. World Health Organization. WHO global strategy on people-centred and integrated health services: interim report. https://apps.who.int/iris/handle/10665/155002 (2015).)

@@

% ***七星山的 idea [2021/03/01]
Generative adversarial networks simulate gene expression and predict perturbations in single cells
3.1. Gene expression in \cite{Ching2018} % deep review
% 環境 -> chromosome accessibility
%  unsupervised deep learning algorithm
generative adversarial networks (GAN) with VAE (variational autoencoder) could simulate environment causing carcinogenesis
%graph convolutional network (from deepchem import GraphConvModel)
deep or graph convolutional network in gene regulation, which includes chromosome accessibility, transcription factor binding site, DNA methylation, and RNA inhibitions (RNAi, microRNA, LncRNA).

%GAN overview
In this paper compares GAN with VAE (variational autoencoder) and RBM (restricted Boltzmann machine) models.
As a new unsupervised learning algorithm framework, generative adversarial networks (GAN) has been favored by more and more researchers, and it has become a research hotspot. GAN is inspired by the two-person zero-sum game theory in game theory. Its unique confrontation training idea can generate high-quality samples and has more powerful feature learning and feature expression ability than traditional machine learning algorithms. At present, GAN has achieved remarkable success in the field of computer vision, especially in the field of sample generation.
%Learning Generative Models of Tissue Organization with Supervised GANs
\cite{Han2018}

%https://www.amazon.com/Deep-Learning-Life-Sciences-Microscopy-ebook/dp/B07QM3G5JM

%book: https://github.com/deepchem/DeepLearningLifeSciences
%https://github.com/deepchem/torchchem

%https://github.com/hindupuravinash/the-gan-zoo

@@

ground truth Y (e.x. patient's survival):\\[0.5cm]
%Y$:\\[0.3cm]
$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n + \epsilon$
gene expression is not an independent $X_mrna$, which is correlated with other $X_body, X_stress, X_inflammation, X_ros$
There is much work to be done in investigating the functional relationships more fully.\cite{Rogers1959}
We wish the TCGA does collect holistic features from their participants in the future.

holistic-enabled Electric healthcare record, EHR
Deep Patient
Under HL7, RESTful
deep learning by EHR-converted FHIR: with holistic observations
=> Level 4 Record-keeping and Data Exchange for the healthcare process
(clinical data)
Allergy, Problem,
Procedure,
CarePlan/Goal,
ServiceRequest,
Family History,
Risk Assessment (APGAR, Glasgow coma scale)
condition: physical, mental
{observations: 
vital signs, 
body temperature, body weight and body height, 
x-ray, 
laboratory data, 
pulse oximetry, 
ECG, EEG, EMG, 
clinical findings (symptom and sign), 
tobacco exposure, family support, cognitive status
pregnancy status
death assertion
HIV, HBV, HCV status
}

%https://mos2718.github.io/FHIRspec/Index.html
%https://www.hl7.org/fhir/patient.html
%https://mos2718.github.io/FHIRspec/Spec/Patient/EHR_Patient.docx

@@

In personal history section of admission note (EHR for inpatient): 
< Habits >
　　　　-Tobacco:
　、　1.5　pack/day for　30　years
　　　　-Alcohol drinking:　Occasionally
　　　　-Betel nuts:　、　nuts/day for　30　years　
　　< Social/Economic Evaluation >
　　　　-Living condition:　Living with family
　　　　-Family support:　Well
　　　　-Main caregiver:　Spouse
　　　　-Financial support:　Sufficient
　　< Psychological Evaluation >
　　　　Stable
　　　　Suspected to be ” abuse ” or ”neglected victim” (if the answer is “Yes“ should be assisted by social workers)　No
　　< Spiritual Assessment >
　　　　Religion:　Nil
　　　　Diagnosed with life-threatening diseases or terminal stage of chronic diseases or cancers?　No
